Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-11-05 robdml.tex
tex

As pushing $\tilde{H}_\mathsf{S}$ towards the upper bound will easily lead to
model collapse, a valid $H_\mathsf{D}$ should be chosen within the interval
$[H_\mathsf{S},2+\gamma]$ without rendering model collapse.
%
Thus, intuitively, $H_\mathsf{D}$ can be the hardness of another benign triplet
(with the same anchor) sampled with a strategy with a higher hardness
expectation, \ie, $E[H_\mathsf{D}] > E[H_\mathsf{S}]$.
%
Or at least the $\text{Var}[H_\mathsf{D}]$ of such sampling strategy has to be
large enough in order to create a notable number of valid adversarial examples.
%
For instance, $H_\mathsf{D}$ can be the hardness of a Semihard~\cite{facenet}
triplet when the \emph{source} triplet is sampled with Random sampler.
%
Predictably, the model performance will be significantly influenced by the
triplet sampling strategies we chose for $H_\mathsf{D}$ in this case.
%
For convenience of further discussion, we denote the hardness of a Random,
Semihard, and Softhard triplets as $\mathcal{R}$,
$\mathcal{M}$, $\mathcal{S}$, respectively.

@@

Even if $H_\mathsf{D}$ is calculated from another triplet harder than the
\emph{source} triplet, the adversarial example may become weak in the late
phase of training.
%
The optimizer aims to reduce the expectation of triplet
loss $E[L_\text{T}]$ to zero as possible over the distribution of sample triplets, and thus
the $E[H]$ of any given triplet will tend to $-\gamma$, reducing the
hardness of adversarial triplets from HM.
%
We speculate this will lower the level of robustness the model could reach.

@@

Thus, $H_\mathsf{D}$ should be (1) close to $H_\mathsf{S}$ in the
early training phase (indicated by a large loss value), and (2) clearly higher
than $H_\mathsf{S}$ in the late training phase (indicated by a small loss
value),
%
in order to automatically balance the training objectives (\ie, performance
on benign examples \emph{v.s.} robustness).
%
We name the two conditions as ``Graduate Adversary'' for 
pseudo-hardness functions, as illustrated in \cref{fig:ga}.

@@

As an example, we propose a ``Linear Gradual Adversary'' (LGA) pseudo-hardness
function that is independent to any benign triplets, incorporating an
assumption that $H_\mathsf{D}$ should remain Semihard~\cite{facenet}, as
follows:
%
\begin{equation}
	%
	g_\mathsf{LGA}(\bar{\ell}_{t-1}) = -\gamma \cdot \bar{\ell}_{t-1} ~ \in
	[-\gamma,0].
	%
\end{equation}
%
And the training objectives will be automatically balanced in
$\text{HM}[H_\mathsf{S},g_\mathsf{LGA}]$.
%
The design choice of LGA is empirically supported by experimental observations.
%
We leave more complex pseudo-hardness functions for future study.
%

>>>>

As pushing $\tilde{H}_\mathsf{S}$ towards the upper bound will easily render
model collapse, a valid $H_\mathsf{D}$ should be chosen within the interval
$[H_\mathsf{S},2]$.
%
Thus, intuitively, $H_\mathsf{D}$ can be the hardness of another benign triplet
(with the same anchor) sampled with a strategy with a higher hardness
expectation, \ie, $E[H_\mathsf{D}] > E[H_\mathsf{S}]$.
%
Or at least the $\text{Var}[H_\mathsf{D}]$ of another benign triplet has to
be large enough (for a small portion of triplets $H_\mathsf{D}>H_\mathsf{S}$)
in order to create a notable number of valid adversarial
examples.
%
For instance, $H_\mathsf{D}$ can be the hardness of a Semihard~\cite{facenet}
triplet when the \emph{source} triplet is sampled with Random sampler.
%
Predictably, the model performance will be significantly influenced by the
triplet sampling strategies we chose for $H_\mathsf{D}$ in this case.
%
For convenience of further discussion, we denote the hardness of a Random,
Semihard, and Softhard triplets as $\mathcal{R}$, $\mathcal{M}$, $\mathcal{S}$,
respectively.

@@

Even if $H_\mathsf{D}$ is calculated from another triplet harder than the
\emph{source} triplet, the adversarial example may become weak in the late
phase of training.
%
The optimizer aims to reduce the expectation of 
loss $E[L_\text{T}]$ towards zero as possible over the distribution of triplets, and thus
the $E[H]$ of any given triplet will tend to $-\gamma$, reducing the
hardness of adversarial triplets from HM as $E[H_\mathsf{D}-H_\mathsf{S}]$
decreases accordingly.
%
Weakened adversarial examples are insufficient for robustness.

@@

Thus, $H_\mathsf{D}$ should lead to (1) relatively weak adversarial examples in
the early training phase (indicated by a large loss value), and (2) relatively
strong adversarial examples in the late training phase (indicated by a small
loss value),
%
in order to automatically balance the training objectives (\ie, performance on
benign examples \emph{v.s.} robustness).
%
A satisfactory pseudo-hardness function is a ``Graduate Adversary'', as
illustrated in \cref{fig:ga}.

@@

As an example, we propose a ``Linear Gradual Adversary'' (LGA) pseudo-hardness
function that is independent to any benign triplets, incorporating an empirical
observation that $H_\mathsf{D}$ should remain Semihard~\cite{facenet}, as
follows:
%
\begin{equation}
	%
	g_\mathsf{LGA}(\bar{\ell}_{t-1}) = -\gamma \cdot \bar{\ell}_{t-1} ~ \in
	[-\gamma,0].
	%
\end{equation}
%
And the training objectives, namely performance on benign examples and
robustness will be automatically balanced in
$\text{HM}[H_\mathsf{S},g_\mathsf{LGA}]$, leading to a better eventual overall
performance.
%
More complicated or non-linear pseudo-hardness functions are left for future study
for sake of simplicity.
%
